### 2018-05-12 msc_diploma_cubesat_galgoczi.tex

>>>>

\subsection{Gamma-ray bursts}

Gamma-ray bursts (GRBs) [1, 2, 3, 4, 5, 6, 7, 8, 9] have been studied already about four decades, nonetheless their origin is still not fully understood. They are one of the most extreme explosive events ever observed. In spite of tremendous effort and numerous observations there are still open questions concerning their detailed physics. GRBs were discovered in years 1967–73 by military Vela satellites and later reported to the scientific community in early seventies [10, 11]. They occur roughly once per day and are characterized by flashes of $\gamma$-rays typically lasting from a fraction of a second to thousands of seconds, typically in the energy range from several keV to several MeV, however $\gamma$-rays of energy >1 GeV associated with GRBs were detected as well. A GRB event outshines any other gamma-ray source on the sky.
It has been found that the duration distribution of their $\gamma$-ray prompt emission is bimodial [12] which suggested that there were two groups of GRBs. Later, based on more observations, it has been confirmed that these two GRB groups were two distinct astrophysical populations [12, 13, 14, 15, 16,17]: I. so called long GRBs (LGRBs) with softer spectrum and with prompt $\gamma$-ray emission $\geqslant$2 s that were identified to be gravitational collapses of massive stars due to their association with type Ic core-collapse supernovae, and II. so called short GRBs (SGRBs) with harder spectrum and with prompt $\gamma$-ray emission $\leqslant$2 s originating in merger of two compact objects such as neutron star - neutron star (NS-NS) or neutron star - black hole (NS-BH) [18].

\begin{figure}[H]
\centering
\includegraphics[width=130.0mm]{images/Gamma_ray_burst.jpg}
\caption{The life cycle of a massive star. When fusion stops the collapsing star releases energy in the form of jets observed as a gamma-ray burst}
\end{figure}

It was proposed that GRBs were created in collisions of highly relativistic outflow of the accelerated jetted matter [19, 20]. In many cases it was found that the prompt $\gamma$-ray emission is followed by longer-lasting afterglow in soft X-ray, optical or radio waves [21, 22] explained as a result of the propagation of a relativistic shockwave through the circumburst medium. The suggestion that GRBs lie at cosmological distances [23] was confirmed by numerous observations and it was found that their redshifts are up to z=9.4 with $\langle$ z $\rangle\approx$0.5 for SGRBs and $\langle$z$\rangle \approx$2.0 for LGRBs [18]. The energy content 51 released in a GRB is $\sim$10 erg.
In some cases of LGRBs there was observed supernova component in the lightcurve or in the spectrum of its afterglow, peaking $\sim$2 weeks after the initial trigger [24, 25, 26]. LGRBs are associated with the brightest regions of galaxies and with the most massive stars [27]. The relativistic outflow of LGRBs has a bulk Lorentz factor of $\Gamma\sim$300 [5]. The afterglow observations indicate that a geometrical beaming of the LGRBs radiation has an opening angle of $\sim$5 [28].

\begin{figure}[H]
\centering
\includegraphics[width=130.0mm]{images/GRB_BATSE_12lightcurves.png}
\caption{The light curves of 12 gamma-ray bursts observed by BATSE. The diversity of the curves is extraordinary. Events with a length from millisecond to a second exist either with smooth or highly variable curves.}
\end{figure}

Observations suggest that SGRBs arise from old astrophysical populations. The beaming angles for SGRBs jets seems to be from $\sim$5 to > 25 [5, 29, 30]. The model that they originate in the merger of two NS has been recently confirmed by the detection of the gravitational wave (GW) signal by LIGO / Virgo collaboration 1, 2. This GW signal was accompanied with a SGRB detected by the Fermi /Gamma-ray Burst Monitor (GBM) and INTEGRAL instruments followed by the observation of the GRB’s afterglow and its host galaxy [31, 32, 33]. In case of SGRBs a kilonova/macronova can be observed. It emits electromagnetic radiation due to the radioactive decay of heavy r-process nuclei that are produced and ejected almost isotropically during the merger process [34].

\subsection{Open questions on gamma-ray bursts}

In spite of great theoretical and observational successes over the last four decades, many key questions remain about the physics of GRBs. For a detailed review see [3, 4, 6, 35]. Some of these key questions are following.

Central engine: it is usually considered that as a result of the core collapse of a massive star or as a result of a merger of two compact objects there is a newborn BH created. Some models also consider that instead of a BH, a rapidly rotating NS with strong surface magnetic field (magnetar), B$\geqslant	$10$^{11}$ T, is formed. Another suggested possibility is a formation of a never observationally confirmed quark star composed of quark matter.
The particle acceleration mechanism in the relativistic shocks and the microphysics of the shocks is a major open issue in the GRB physics. 
GRB jet composition: it is still debated if the GRB outflow is baryonic or magnetic field (Poynting flux) dominated. The jet composition is tied to whether GRBs can accelerate protons to ultra-high energies and produce Ultra-High Energy Cosmic Rays (UHECRs) and high energy neutrinos. The models which expect GRBs to be sources of UHECR and high energy neutrinos invoke a baryon-dominated outflow.
Radiation mechanism: the prompt emission of GRBs is non-thermal, smoothly-joint broken-power-law called “Band” function. The emission is required to be produced by a population of particles, e.g. leptons, with a power law distribution of their energies. The main assumed mechanisms are synchrotron radiation, Synchrotron Self-Compton radiation, and Compton up-scattering of thermal photons however it is difficult to identify the correct radiation mechanism with the current available
data.
First stars (Pop III): what is the expected signature of GRBs originating in the death of the first metal-free massive stars at high redshift? It was estimated that the mass of the BH at the center of the collapsar of Pop III star is a factor of ten larger and Pop III GRBs may thus tend to have larger durations than those of ordinary Pop I and Pop II GRBs [36].
Classification: there are cases of GRBs which do not exactly fit in the usual short/long classification.
For example the ultra-long GRBs or short GRBs with soft extended emission. Next question is what is
the nature of the precursors in some of the GRB light curves?
Recent detection of the electromagnetic (EM) signal of the short GRB 170817A [37] associated with a GW signal GW170817 produced by a binary neutron star (BNS) merger [32] marks a milestone in the multi-messenger research of GRBs and means a great importance of the efforts of searching for EM counterparts to GW sources with spaceborn instruments. The discovery was made on August 17, 2017 at 12:41:06 UTC with the Fermi/GBM which detected and triggered on the short gamma-ray burst GRB 170817A. Approximately 1.7 prior to this GRB, LIGO-Virgo detector network observed a GW signal GW170817 at 12:41:04.4 UTC which originated from a BNS inspiral. The source was localized 9/89within a sky region of 28 deg 2 (90 \% probability), which was the most precisely localized GW signal yet. An extensive observing campaign was launched across the electromagnetic spectrum [33] leading to the discovery of a bright optical transient in NGC 4993. Observations in $\gamma$-ray, X-ray, and ultraviolet (UV) wavelengths include also INTEGRAL [38], Chandra X-ray Observatory [39], Swift and Nuclear Spectroscopic Telescope ARray (NuSTAR) were performed and a blue kilonova associated with this event was detected [40].
Some models predict also EM signal from BBH coalescence, for example [41, 42]. More detections of EM counterpart to GW sources and kilonovae are indeed essential for better understanding of these phenomena and this proposed mission has a great potential to bring new and valuable observational data.
After the upgrades of LIGO, following the finished observing run O2, there will be O3 run scheduled to start in 2018. Further observing runs are expected after next improve of the sensitivity. Designed sensitivity is aimed to be achieved in 2021. The Kamioka Gravitational Wave Detector (KAGRA) 3 is expected to start the first cryogenic operation in 2018, and the observing runs with a full interferometer are expected in 2020s [43]. Therefore, more GW detections are expected during the operation of this proposed mission.

\subsection{Particle detectors in space}

Various detector types has been used or are planned for space instruments for example gas-filled detectors like proportional counters and Geiger counters; scintillation detectors like inorganic crystal scintillators, e.g. NaI(Tl) or CsI(Tl), and plastic scintillators; semiconductor devices like, e.g. cryogenically cooled germanium, Cadmium Zinc Telluride (CZT) or Silicon Drift Detector (SDD) technique. Other types of detectors used in space missions are Compton scatter telescopes or pair production telescopes. Table 3 lists detector technologies capable of detecting GRBs in hard X-ray and gamma-ray range employed in previous and current major instruments. For an extensive review of hard X-ray / soft gamma-ray experiments and missions see [47].

\subsection{Current missions observing gamma-ray bursts}

Several spaceborn instruments and ground-based observatories are currently operating or are planed in
order to bring more detailed observations of GRBs to better understand their nature. Their overview
and comparison with the required specification on our proposed mission is given in following
subsections.
Our project will benefit from large field of view (FoV) and high localization accuracy. It will provide
full-sky FoV and due to the advantage of using multiple satellites the occulted area by the Earth will
be eliminated. The expected localization accuracy will be from $\sim$10' to degrees depending on the flux and duration of a GRB. This is not achieved by any of the current GRB missions. In the current missions either the large FoV or precise localization accuracy is accomplished. The large FoV instruments, e.g. INTEGRAL/SPI-ACS or Fermi/GBM, have either no localization capability or the localization accuracy is of the order of several degrees. On the other hand the current instruments providing localization accuracy of the order of arcmin have limited FoV, e.g. Swift/BAT or
AGILE/Super-AGILE.
Another advantage of our proposed mission is that the cost is only a fraction of the cost of a big
satellite missions, e.g. like SVOM, and the concept of multiple cubesats allows higher flexibility in
terms of extending the fleet by more satellites as well as allows replenishment of failed satellites.

Fermi 7 [44, 45, 66] was launched in 2008 and the main instruments are the Large Area Telescope (LAT) and the Gamma-ray Burst Monitor. 
Neil Gehrels Swift Observatory (Swift) 8 [64, 67] was launched in 2005 and the main instruments are
Burst Alert Telescope (BAT), X-ray Telescope (XRT), and UV/Optical Telescope (UVOT).
INTErnational Gamma-Ray Astrophysics Laboratory (INTEGRAL) 9 [57, 58, 68, 69] was launched in
2002 Among several instruments on board INTEGRAL the ones which detect most of GRBs are the Anti-Coincidence Shield of the SPectrometer of INTEGRAL (SPI-ACS) and the Integral Soft Gamma-Ray Imager (ISGRI) detector layer of the Imager on Board the Integral Satellite (IBIS).
Wind/Konus 10 [70] was launched in 1994 and its still active GRB instrument is Konus.
Astro-Rivelatore Gamma a Immagini Leggero (AGILE) 11 /Super-AGILE [48, 71] was launched in
2007 and its Hard X-ray Imaging Detector (Super-AGILE) detected several GRBs.
CALorimetric Electron Telescope (CALET) 12 [52, 72, 73] / Gamma-Ray Burst Monitor (CGBM), was launched in 2015 and it is placed on the International Space Station (ISS).

Monitor of All-sky X-ray Image (MAXI) 13 [60, 74 ], started nominal observation in 2009. Its instrument Gas Slit Camera (GSC) detects GRBs. The instrument is placed on ISS.
Gamma-ray Burst Polarimeter POLAR 14 [61, 75], was launched in 2016 and is dedicated to the measurement of GRB polarization.
Lomonosov 15 /BDRG [59, 76], was launched in 2016. The BDRG instrument on board detected several GRBs.
The Hard X-ray Modulation Telescope (Insight-HXMT) 16 [56, 77], was launched in 2017. It has on board the high energy X-ray telescope (HE), the medium energy X-ray telescope, and the low energy X-ray telescope.
ASTROSAT 17 [49, 78, 79], was launched in 2015 and its Cadmium Zinc Telluride Imager (CZTI) has already detected many GRBs.
Reuven Ramaty High Energy Solar Spectroscopic Imager (RHESSI) 18 [62, 80, 81], was launched in 2002. It is designed to study hard x-ray and gamma-ray emission from solar flares, however it is also efficient instrument to detect non-solar gamma-ray events like GRBs.
The energy range, FoV and localization accuracy for the instruments of the current missions which observe GRBs is summarized in Table 4. Figure 3 shows the FoV vs. localization accuracy of the current instruments which detect GRBs. For comparison the specification of this project is shown as
well.

\subsection{Currently operating GRB collection and alert networks}

Our project can potentially contribute to some of the currently operating GRB collection and alertnetworks such as the following ones:
The InterPlanetary Network (IPN) 27 [97, 98] which derives the positions of fast gamma-ray transients of all kinds by triangulation. Numerous spacecraft and instruments participate in the network at the Earth’s orbit as well as in the interplanetary space. It detects about 350/year.
The Gamma-ray Coordinates Network (Transient Astronomy Network) (GCN/TAN) 28 provides information about GRBs in real-time to the world community. The network provides the real-time (and near real-time) distribution of GRB locations, images, spectra, and lightcurves detected by various spacecraft and the distribution of follow-up observation reports.
The Global Relay of Observatories Watching Transients Happen (GROWTH) 29 is an international scientific collaborative project studying the astronomical transients.

\subsection{Current GRB Follow-up observations}

This cubesat project will provide high precision localization accuracy of GRBs allowing efficient follow-up observations by existing ground based observatories. There are several observatories world-wide to perform photometric or spectroscopic observations of GRB afterglows, for example:
The Mobile Astronomical System of TElescope Robots (MASTER) 30 [99] is a global network of robotic telescopes for very fast positioning and follow up of GRBs.
The Burst Optical Observer and Transient Exploring System (BOOTES) network of robotic telescopes for follow up observations of GRBs. 31 [100] is a world wide The Robotic Optical Transient Search Experiment (ROTSE) 32 [101] are telescopes which operate at
sites around the world and follow GRBs.
Pi of the Sky 33 [102] is a system of detectors designed for continuous observation of night sky looking for optical flashes of astrophysical origin, in particular for GRBs.
Several other telescopes and observatories which provide measurements in optical, infra-red or radio bands are also, e.g.: Gamma-Ray Burst Optical/Near-Infrared Detector (GROND) 34 , the Palomar 60 inch telescope 35 , Rapid Eye Mount (REM) 36 , Livermore Optical Transient Imaging System (Super- LOTIS) 37 , Télescopes à Action Rapide pour les Objets Transitoires (TAROT) 38 , Gran Telescopio Canarias (GTC) 39 , ESO Very Large Telescope (VLT) 40 , Nordic Optical Telescope (NOT) 41 , Observatorio Astronómico Nacional on the Sierra de San Pedro Mártir 42 , Cerro Tololo Inter-American Observatory (CTIO) 43 , Telescopio Nazionale Galileo (TNG) 44 , Gao-Mei-Gu (GMG) station of Yunnan Observatories 45 , Observatorio de Sierra Nevada (OSN) 46 , United Kingdom Infrared Telescope (UKIRT) 47 , Very Large Array (VLA) 48 , Atacama Large Millimeter/Submillimeter Array (ALMA) 49 , James Clerk Maxwell Telescope 50 , Arcminute Microkelvin Imager which robotically triggers on Swift GRBs (AMI-GRB) 51 .
